---
title: "Presentation - Shiny App"
author: "Dario J C"
date: "25/04/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(WikidataQueryServiceR)
library(stringr)
library(readr)
library(dplyr)
library(tidyr)


# Query Wikipedia for scientists
scraped_raw <- query_wikidata('SELECT ?scientist ?scientistLabel ?birthdate ?countryLabel ?coord WHERE {
  ?scientist ?x1 wd:Q901;
  wdt:P569 ?birthdate;
  wdt:P19 ?birthplace;
             wdt:P27 ?country.
 
  ?birthplace wdt:P625 ?coord
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
ORDER BY DESC(?birthdate)')

# Query Wikipedia for fictional scientists to remove
fictional_scrape_raw <- query_wikidata('SELECT ?scientist ?scientistLabel ?birthdate ?countryLabel ?narrative ?coord WHERE {
  ?scientist ?x1 wd:Q901;
  wdt:P569 ?birthdate;
  wdt:P19 ?birthplace;
             wdt:P1080 ?narrative;
             wdt:P27 ?country.
 
  ?birthplace wdt:P625 ?coord
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
ORDER BY DESC(?birthdate)')


# Clean the query results
table_clean <- scraped_raw %>%
  mutate( coord = str_replace_all(string = coord,
                                  pattern = "(Point|\\(|\\))",
                                  replacement = ""),
          birthdate = as.Date(birthdate)
  ) %>%
  separate( col = coord,
            into = c("long", "lat"),
            sep = " ",
            remove = TRUE) %>%
  mutate( long = parse_number(long),
          lat = parse_number(lat)) %>%
  rename(country = countryLabel,
         link = scientist,
         scientist = scientistLabel) %>%
  filter(link != fictional_scrape_raw$scientist &
           str_count(scientist, pattern = " ") >= 1
  ) %>%
  select(scientist,
         birthdate,
         country,
         long,
         lat,
         link)


```

## Scientists & Their Birth Places

The goal of this app is to let the user explore where scientists have been born. 

* The Data is directly queried from Wikipedia ([Wikidata](https://www.wikidata.org/wiki/Wikidata:Main_Page)).
* The code can be found on [github](https://github.com/dario-j-c/Developing-Data-Products-Week-4).
* The app can be found on [shiny](https://dario-j-c.shinyapps.io/scientists_location/).

## Data Used

The r package, [WikidataQueryServiceR](https://github.com/wikimedia/WikidataQueryServiceR) was used to query and get the raw data. That data was minimally cleaned until we got the following.

```{r summary, echo = TRUE}
head(table_clean) %>%
  gt::gt()

```


## Method

The app allows the user to:

* Automatically query Wikidata about all their listed scientists
* Filter the generated table by date of birth and / or place of citizenship (country)
* Explore (before or after filtering) the generated map and table listing
  * The scientists' names
  * The location of their birth (this is not always the same as place of citizenship)

## Conclusion

Enjoy & Thank You
